What is an Ancestral Recombination Graph?

### 2022-02-14 paper.tex
First stab at ancestry resolution text

\section*{Ancestry resolution and simplification}

\begin{figure}
\centering
\vspace{5em}
\caption{\label{fig-ancestry-resolution}
}
\end{figure}

>>>>

\section*{Ancestry resolution and simplification}
% Note: we want some terminology here to be precise about what we're
% talking about - perhaps a G-ARG for genome-ARG vs E-ARG for event?
% Just using ARG for now, can clean up in another pass.
The definition of an ARG as a data structure in terms of
individuals' monoploid genomes
and their direct patterns of inheritance allows us to describe arbitarily
complex relatedness structures, free from the limitations imposed by
coalescent approximations.
An ARG is an intrinsically retrospective
structure, defined with respect to the ancestry of
a set of sampled genomes.
The graph structure (the nodes and their connectivity through edges)
captures a great deal of information about this ancestry,
but on its own it cannot tell us about the precise relationships between
the samples along the genome: do to this, we must construct the
local (or ``marginal'') trees.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{illustrations/ARG_edge_annotations}
\caption{\label{fig-ancestry-resolution}
Implicit vs explicit edge annotations. (a) An eARG equivalent to the encoding
described by Griffiths can be created by setting all edges to span the full
genome, apart from edges between two recombining nodes and their recombinant
child (edge spans in bold). (b) The sample-resolved eARG, where the Hudson
simplification algorithm has shortened edge spans so that they encapsulate only
the upwards transmission of genetic material from the samples A, B, and C;
plotted line widths are proportional to the genomic span of each edge. The
example ARG used here is taken from \citet[][fig. 1]{wiuf1999recombination}.
}
\end{figure}

% This is weak. Should we give a better description of the algorithm?
% Not clear yet, and don't want to spend too much time on it if it's
% unhelpful.
Given a set of edges $(u, v, I)$ where node $u$ descends from $v$
over the set of genomic intervals $I$, and a set of sample nodes $S$
it is straightforward to construct the local tree at position $x$.
Starting from the sample nodes in $S$, traverse upwards through the
graph, such that when we are at a node $u$ we follow the edge
$(u, v, I)$ where $x \in I$. This is essentially the same procedure as
described by Griffiths~\citep{griffiths1997ancestral}.

Recovering the tree at a single point along the genome, although
a fundamental operation, is not the only operation that we may wish
to perform on an ARG. We can make the operation of efficiently
constructing \emph{all} trees along the genome much more efficient,
and gain insight into the structure of the ARG,
if we perform an additional step and \emph{resolve} the ancestral
material in the ARG with respect to the samples.

As we saw in the section X, the key to efficiently accessing the
results of the ARG stochastic process is to reason about
the flow of ancestral material through the graph using Hudson's
algorithm. The same basic process allows us to determine the
exact intervals of ancestral material that are flowing through
each edge in the graph. Concretely, given a set of edges
$E = \{(u, v, I)\}$ where $u$ is the child, $v$ is the parent
and $I$ is the set of genomic intervals passed directly from
$v$ to $u$, the resolution process with respect to a set of
samples $S$ produces a set of edges $E' = \{(u, v, I)\}$
where $I$ is the set of genomic intervals over which there is a
% weak and not correct - how do we say this properly?
path from at least one element of $S$ to $u$,
and $u$ is not an MRCA.

This process is illustrated in Fig~\ref{fig-ancestry-resolution}.
